% Copyright (c) 2021 Robert Carnell

\documentclass[11pt,letter]{book}

%\usepackage[T1]{fontenc}
\usepackage{fontspec}

% Fonts added for the book
%\usepackage[bitstream-charter]{mathdesign} % use the bitstream-charter True Type font

% ---- Gramps Packages ----
%\usepackage[latin1]{inputenc}%
%\usepackage[latin1,utf8]{inputenc}% need to use XeLaTex
\usepackage{graphicx}% Extended graphics support
\usepackage{longtable}% For multi-page tables
\usepackage{calc}% For some calculations
\usepackage{ifthen}% For table width calculations
\usepackage{ragged2e}% For left aligning with hyphenation
\usepackage{wrapfig}% wrap pictures in text

% Packages added for the book
\usepackage[all]{genealogytree} % genealogy-tree package for trees
\usepackage[hidelinks]{hyperref} % to break long URL lines, and make TOC clickable
\usepackage{endnotes} % create endnotes
\usepackage{etoolbox} % modify endnotes
% endnote modifications
\renewcommand{\notesname}{Endnotes}
\patchcmd{\theendnotes}
  {\makeatletter}
  {\makeatletter\renewcommand\makeenmark{\theenmark.\space}}
  {}{}

% Begin document
\title{\bf Carnell-Rettgers Family Genealogy \\
       \large Volume V - News Articles and Obituaries}
\author{Robert Carnell}
\date{\today}

\begin{document}
\frontmatter
\maketitle
\clearpage

\begingroup
\parindent 0pt
\parskip
\baselineskip
Copyright \textcopyright{} 2021 Robert Carnell

All rights reserved.  This book or any portion thereof may not be reproduced or used in any manner whatsoever with the express written permission of the author except for the use or brief quotations in a book review or genealogical research.
\endgroup
\clearpage

\tableofcontents
\mainmatter

\chapter{Introduction}

The newspaper articles, obituaries, death notices, and marriage notices contained in this volume have been essential in bringing together the family's genealogy.  They have been an invaluable source of accurate family information which is otherwise difficult to obtain.  On more than one occasion, information in an obituary helped to correct errors in many historical references which relied on gravestone and memories.

The articles in this volume written after 1922 may still covered by the copyright of the publishing newspaper.  The transcripts were obtained and copied as a fair use of the articles for research purposes.  Each article is accompanied by a citation.  No permission has been obtained to copy or distribute any article in this volume.

\chapter{News Articles}
\renewcommand{\thesection}{\arabic{section}}

\input{../tex_output/obituaries.tex}

\end{document}
